---
title: "EDA"
author: "Connor Birkhold"
date: "1/6/2020"
output: html_document
---

#### The goal of this EDA is to explore the dataset while validating its accuracy and looking for ways to reduce the size of the data without compromising its integrity.


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 5)
```

```{r include=FALSE}
library(data.table)
library(tidyverse)
library(lubridate)
```

#### Download the data
```{r}
url <- "https://data.sfgov.org/resource/wg3w-h783.csv"

sfCrime <- fread(url, encoding = 'UTF-8')

str(sfCrime)
```

#### Identify missing values
```{r}
# If there are blank values convert them to NAs
sfCrime[sfCrime == "", ] <- NA

# Calculate the percentage of NAs by column
pct.NA <- function(x){
    sum(is.na(x))/length(x) * 100
}

pctMissVal <- as.data.frame(lapply(sfCrime, pct.NA))
pctMissVal <- gather(pctMissVal, key = 'c_name', value = 'pct_na') %>% 
    arrange(desc(pct_na))

# List columns with percentage of NAs from greatest to least
print(pctMissVal[1:10, ])
```

#### With a high percentage of NAs I want to know more about what makes up the @computed_region columns? Are they useful? Are they candidates for imputation?
```{r results='hide'}
# Let's look into a @computed_region column with less NAs
table(sfCrime$`:@computed_region_qgnn_b9vv`)

# Maybe it is a coding for the district or neighborhood?
table(sfCrime$`:@computed_region_qgnn_b9vv`, sfCrime$police_district)
table(sfCrime$`:@computed_region_qgnn_b9vv`, sfCrime$analysis_neighborhood)

# Not sure of their significance at this point in time. Removing all @computed_region columns because large amount of NAs and don't think they are significant
colNm <- grep("@computed_region", names(sfCrime), value = TRUE)
sfCrime <- sfCrime[, !colNm, with = FALSE]
```

#### What are my percentage of NAs by column with @computed_region removed?
```{r}
# the "@" was converted to "X.." in pctMissVal so running grep again
rowNm <- grep("X..computed_region", pctMissVal$c_name, value = TRUE)

'%ni%' <- Negate("%in%") # creat not in function

pctMissVal <- pctMissVal %>% filter(c_name %ni% rowNm)

print(pctMissVal[1:10,])
```

#### Removing columns filed_online and cad_number due to high percentage of NAs
```{r}
sfCrime <- sfCrime[, -c('filed_online',
                        'cad_number')]
```

#### Cleaning up date columns
```{r}
# Only interested in the dates right now. Cleaning those up and removing times
sfCrime <- sfCrime[, 
                   `:=`(incident_date = str_sub(incident_date, 1, 10),
                        report_date = str_sub(report_datetime, 1, 10))] %>% 
    .[,
      `:=`(incident_date = ymd(incident_date),
           report_date = ymd(report_date))]

# excluding datetime and time
sfCrime <- sfCrime[, -c("incident_datetime", 
                        "report_datetime",
                        "incident_time")]
```

#### According to the documentation provided with the dataset each row represents an incident. Let's validate which column or columns support this
```{r}
# row_id
uniqueN(sfCrime$row_id) == nrow(sfCrime); uniqueN(sfCrime$row_id)

# incident_id
uniqueN(sfCrime$incident_id) == nrow(sfCrime); uniqueN(sfCrime$incident_id)

# incident_number
uniqueN(sfCrime$incident_number) == nrow(sfCrime); uniqueN(sfCrime$incident_number)

# combo of incident_id, incident_number, and incident_code
uniqueN(paste(sfCrime$incident_id, 
              sfCrime$incident_number, 
              sfCrime$incident_code, sep = "")) == nrow(sfCrime); uniqueN(paste(sfCrime$incident_id, 
              sfCrime$incident_number, 
              sfCrime$incident_code, sep = ""))

# setting up a unique column with the combination of incident_id, incident_number, and incident_code
sfCrime <- sfCrime[, incident_id_nbr_cd := paste(sfCrime$incident_id, 
              sfCrime$incident_number, 
              sfCrime$incident_code, sep = "-")]

# remove incident_id, incident_number, incident_code, and row_id. Reducing size of data
sfCrime <- sfCrime[, - c("incident_id",
                         "incident_number",
                         "incident_code",
                         "row_id")]

```

#### Location information
- Can we reduce the location information to only police_district, latitude, and longitude?
- Removing incidents without any location detail removed 69 observations or 6.9% of data
```{r}
# remove "point" because it not required for our intentions
sfCrime <- sfCrime[, - "point"]

# Remove incidents without any location detail
sfCrime <- sfCrime[, filt_ind := ifelse(is.na(latitude) & 
                                            is.na(intersection) &
                                            is.na(longitude) &
                                            is.na(supervisor_district) & 
                                            is.na(analysis_neighborhood), 1, 0)] %>% 
    .[filt_ind == 0, ] %>% 
    .[, - "filt_ind"]

# Referencing the pctMissVal table I can see that police district does not have any NAs and could be used a general location reference for incidents without a latitude and longitude
table(sfCrime$police_district); uniqueN(sfCrime$police_district)

# For now, excluding incidents without a lat/long.
# This step did not remove any additional data
sfCrime <- sfCrime[!is.na(latitude) & !is.na(longitude), ]


# How are police_district and supervisor_district connected?
# Doesn't appear to be a clean connection. Removing supervisor_district for now
table(sfCrime$police_district, sfCrime$supervisor_district)
sfCrime <- sfCrime[, - "supervisor_district"]

# Let's exclude anything with police_district equal to "Out of SF"
sfCrime <- sfCrime[police_district != "Out of SF", ]


# Excluding intersection for now. cnn is also related to intersection, excluding for now
sfCrime <- sfCrime[, - c("intersection",
                         "cnn")]

```

#### Description of incidents
```{r}

# resolution is a column that might infer there will be more or less incidents in the future. Let's clean up resolution. Setting it up to expand this column into indicator columns later when fitting models
sfCrime <- sfCrime[, resolution := tolower(resolution)]
sfCrime <- sfCrime[, resolution := str_replace_all(resolution, " ", "_")]


# report_type_description is another column that might infer there will be more or less incidents in the future. Let's clean up report_type_description. Setting it up to expand this column into indicator columns later when fitting models
sfCrime <- sfCrime[, report_type_description := tolower(report_type_description)]
sfCrime <- sfCrime[, report_type_description := str_replace_all(report_type_description, " ", "_")]

# incident_category, incident_subcategory, and incident_description hold a lot of good detail. Potential to extract key words/phrases for model building (NLP). For now, let's clean up the column values


# Removing report_type_code
sfCrime <- sfCrime[, - "report_type_code"]

```

#### Re-organizing column order
```{r}
# Add an incident_cnt column
sfCrime$incident_cnt <- 1

# Re-order columns
sfCrime <- sfCrime[, c("incident_id_nbr_cd",
                       "incident_date",
                       "incident_day_of_week",
                       "incident_year",
                       "report_date",
                       "police_district",
                       "analysis_neighborhood",
                       "latitude",
                       "longitude",
                       "report_type_description",
                       "incident_category",
                       "incident_subcategory",
                       "incident_description",
                       "resolution",
                       "incident_cnt")]

```

#### Distributions and summary statisitcs
```{r}
# What is the total nbr of incidents per year?
sfCrime[, .(incidents = sum(incident_cnt)),
        by = incident_year]

# What is the avg nbr of incidents per day?
sum(sfCrime$incident_cnt)/uniqueN(sfCrime$incident_date)

# What is the avg nbr of incidents per day by district?
sfCrime[, .(avg_incidents = sum(incident_cnt)/uniqueN(incident_date)),
        by = police_district]

# What is the percentage of incidents by district?
# Central, Northern, and Southern have the highest percentage of incidents. Is that potentially due to the size of the district?
# Could be beneficial to bring in population/demographic data 
sfCrime[, round(prop.table(table(incident_cnt, police_district), 1), 2)]

# What is the percentage of incidents by day of week?
# Most incidents happen on Friday and Saturday (makes sense)
sfCrime[, round(prop.table(table(incident_cnt, incident_day_of_week), 1), 2)]

# What is the percentage of incidents by resolution?
# resolution doesn't appear to be a good indicator column. Removing from data
sfCrime[, round(prop.table(table(incident_cnt, resolution), 1), 2)]
sfCrime <- sfCrime[, - "resolution"]

```

```{r}
# Incidents by date
byDate <- sfCrime[, .(cnt = sum(incident_cnt)),
                  by = incident_date]

byDate <- byDate[order(rank(incident_date)),]

plot(byDate$incident_date, byDate$cnt)
# the data looks incomplete. There are gaps when reports are coming in and then they stop
# and then start again. The dates that look valid (i.e. reports are frequent) are 2018-10-05 to 2018-10-08 and 2019-09-29 to 2019-10-06 
```

```{r}
# Time between incidents

```

